\section{Rendering - Theory}\label{sec:rendering---theory}

\subsection{Raytracing}\label{subsec:raytracing}

\subsubsection{Overview}\label{subsubsec:rt-overview}

\subsubsection{Acceleration Structures}\label{subsubsec:acceleration-structures}

Acceleration structures are essential for reducing the computational complexity of ray-tracing
from $\mathcal{O}(N \cdot M)$ to a logarithmic scale.
They utilize spatial or object subdivision to quickly cull geometry that a ray cannot intersect.
This section is based on lecture 1 of a course from the TU Wien Faculty of Informatics,
Institute of Visual Computing \& Human-Centered Technology,
Research Unit of Computer Graphics~\cite{tuwien_rendering_2020}.

\paragraph{Bounding Volumes}\label{par:bounding-volumes}

Bounding volumes simplify complex geometry into basic shapes
that are computationally cheap to test against rays.
Common types include:

\begin{itemize}
    \item \textbf{\gls{aabb}}: {
        Defined by minimum and maximum extents along the coordinate axes ($x, y, z$).
        They are fast to build and test (using the slab method) but may fit underlying geometry poorly
        if the object is rotated, leading to more false positives during traversal.
    }
    \item \textbf{\gls{obb}}: {
        Similar to \glspl{aabb} but aligned with the object's principal axes rather than the world axes.
        They provide a much tighter fit, reducing the number of false positive intersections,
        but are more expensive to build and test.
    }
    \item \textbf{Bounding Sphere}: {
        Defined by a center point and a radius.
        They are rotationally invariant and offer the fastest intersection test
        (procedural spheres work well with way math as described in section~\ref{subsubsec:rt-overview}),
        but often provide a bad fit for complex objects because the spheres will inevitably overlap.
    }
\end{itemize}

\paragraph{Spatial vs. Object Subdivision}\label{par:subdivision-types}

\begin{itemize}
    \item \textbf{Spatial Subdivision}:
        The space itself is divided into cells like Grids, Octrees, or \glspl{kdtree}.
        Objects that span boundaries must be referenced in multiple nodes, potentially increasing memory usage.
    \item \textbf{Object Subdivision}:
        The set of objects is partitioned into a hierarchy, most notably the \gls{bvh}.
        A \gls{bvh} is a tree structure where each node represents a bounding volume
        that encompasses all geometry in its children.
        Unlike spatial subdivision, each geometric primitive belongs to exactly one leaf node.
        Bounding volumes of sibling nodes can overlap,
        but memory consumption is generally lower and predictable ($2N-1$ nodes for $N$ primitives).
\end{itemize}

\paragraph{Building Strategies}\label{par:building-strategies}

Top-down construction is a common approach, particularly for \glspl{bvh}.
The set of primitives is recursively split into child nodes
until a leaf criterion (e.g., max triangles per node) is met.
The quality of the tree depends heavily on the split strategy:

\begin{itemize}
    \item \textbf{Spatial Median}:
        Splits the node at the spatial center of its bounding box.
        Simple but can result in unbalanced trees if geometry is clustered.
    \item \textbf{Object Median}:
        Splits the list of objects into two equal halves based on their centroids.
        Ensures balanced trees but may result in large overlapping bounding boxes.
    \item \textbf{\gls{sah}}:
        A greedy algorithm that attempts to minimize the expected traversal cost.
        It evaluates potential splits based on the probability of a ray hitting a child node
        (proportional to its surface area) and the cost of traversing it.
        \[
        Cost(b) = SA_L(b) \cdot N_L(b) + SA_R(b) \cdot N_R(b)
        \]
        where $SA$ is the surface area and $N$ is the number of primitives.
\end{itemize}

\paragraph{Comparison}\label{par:comparison}

The choice of acceleration structure involves trade-offs between build time, memory usage, and traversal speed,
as shown in Table~\ref{tab:comparison-of-as}.

\begin{table*}[h]
    \centering
    \caption{Comparison of Spatial Acceleration Structures\cite{tuwien_rendering_2020}}
    \label{tab:comparison-of-as}
    \begin{tabular}{lccc}
        \toprule
        \textbf{Structure} & \textbf{Memory} & \textbf{Build Time} & \textbf{Traversal Quality} \\
        \midrule
        Regular Grid & Low -- High & Low & Poor (non-uniform) \\
        Quad/Octree & Low -- High & Low & Good \\
        \gls{kdtree} & Low -- High & Low -- High & Good -- Excellent \\
        \gls{bvh} & Low & Low -- High & Good -- Excellent \\
        \bottomrule
    \end{tabular}
\end{table*}

\subsection{The Rendering Equation}\label{subsec:rendering-equation}

The fundamental mathematical model for simulating light in physically-based rendering is the Rendering Equation,
introduced by Kajiya in 1986\cite{10.1145/15886.15902}.
It describes the equilibrium distribution of radiance in a scene.
This section is based on lecture 2 of the TU Wien Rendering course~\cite{tuwien_rendering_2020}.

\subsubsection{Radiometry and Photometry}\label{subsubsec:radiometry-photometry}

It is important to understand the difference between radiometry (physical measurement of \gls{em} radiation)
and photometry (light as it is perceived by the human eye).
Both systems define their own set of units.
The conversion between the two is defined by the spectral luminous efficiency function $V(\lambda)$,
which peaks at approximately $555nm$ (green).
This is why colors like green and yellow appear brighter to the human eye compared to colors like blue.

Table~\ref{tab:radiometric-quantities} shows the key radiometric quantities,
and Table~\ref{tab:photometric-quantities} lists their photometric equivalents.
Note that while rendering calculations are performed using radiometric units (specifically Radiance),
output devices (displays) and human perception operate in the photometric domain (Luminance).

\begin{table*}
    \centering
    \caption{Radiometric Quantities (Physical)~\cite{tuwien_rendering_2020}}
    \label{tab:radiometric-quantities}
    \begin{tabular}{lccl}
        \toprule
        \textbf{Quantity} & \textbf{Symbol} & \textbf{Unit} & \textbf{Description} \\
        \midrule
        Radiant Energy & $Q_e$ & Joule [$\mathrm{J}$] & Total energy of radiation \\
        Radiant Flux (Power) & $\Phi_e$ & Watt [$\mathrm{W}$] & Energy per unit time ($\mathrm{J/s}$) \\
        Radiant Intensity & $I_e$ & Watt/sr [$\mathrm{W/sr}$] & Flux per unit solid angle \\
        Irradiance & $E_e$ & Watt/$\mathrm{m}^2$ [$\mathrm{W/m^2}$] & Incident flux per unit area \\
        Radiance & $L_e$ & Watt/($\mathrm{sr} \cdot \mathrm{m}^2$) & Flux per unit projected area per unit solid angle \\
        \bottomrule
    \end{tabular}
\end{table*}

\begin{table*}
    \centering
    \caption{Photometric Quantities (Perceptual)~\cite{tuwien_rendering_2020}}
    \label{tab:photometric-quantities}
    \begin{tabular}{lccl}
        \toprule
        \textbf{Quantity} & \textbf{Symbol} & \textbf{Unit} & \textbf{Description} \\
        \midrule
        Luminous Energy & $Q_v$ & Talbot [$\mathrm{T}$] & Perceived energy of light \\
        Luminous Flux & $\Phi_v$ & Lumen [$\mathrm{lm}$] & Perceived power \\
        Luminous Intensity & $I_v$ & Candela [$\mathrm{cd}$] & Flux per unit solid angle ($\mathrm{lm/sr}$) \\
        Illuminance & $E_v$ & Lux [$\mathrm{lx}$] & Incident flux per unit area ($\mathrm{lm/m^2}$) \\
        Luminance & $L_v$ & Nit [$\mathrm{cd/m^2}$] & Perceived brightness of a surface \\
        \bottomrule
    \end{tabular}
\end{table*}

\subsubsection{The Equation}\label{subsubsec:the-equation}

The Rendering Equation expresses the total spectral radiance $L_o$
leaving a point $x$ in direction $\omega_o$ as the sum
of emitted radiance $L_e$ and reflected radiance:

\begin{align}
    L_o(x, \omega_o) &= L_e(x, \omega_o) + \notag \\
    & \int_{\Omega} f_r(x, \omega_i, \omega_o) L_i(x, \omega_i) (\omega_i \cdot n) d\omega_i
    \label{eq:rendering-equation}
\end{align}

where:
\begin{itemize}
    \item $L_o(x, \omega_o)$ is the total outgoing radiance.
    \item $L_e(x, \omega_o)$ is the emitted radiance (for light sources).
    \item $\Omega$ is the unit hemisphere centered at normal $n$ containing all incoming directions $\omega_i$.
    \item $f_r(x, \omega_i, \omega_o)$ is the \gls{brdf}, defining how light is scattered.
    \item $L_i(x, \omega_i)$ is the incoming radiance from direction $\omega_i$.
    \item $(\omega_i \cdot n)$ is the geometry term (Lambert's cosine law),
        accounting for the spreading of light over the surface area (often written as $\cos\theta_i$).
\end{itemize}

\paragraph{Usage of Radiance ($L$)}

Radiance ($L$) is the central quantity in rendering for several reasons:
\begin{enumerate}
    \item \textbf{Invariance}: Radiance is constant along a ray in a vacuum.
    This implies $L_i(x, \omega) = L_o(y, -\omega)$ where $y$ is the visible surface point.
    This property enables ray tracing: to find the radiance reaching the camera,
    we simply trace a ray into the scene and calculate the radiance leaving the intersected surface.
    \item \textbf{Sensor Response}: Camera sensors and the human eye respond to radiance
    (integrated over the pixel area and solid angle).
    \item \textbf{Light Transport}: It describes the flow of light energy through space.
\end{enumerate}

The recursive nature of $L_i$ (which is the $L_o$ from another point)
makes this integral equation difficult to solve analytically.
Therefore, most implementations use numerical methods like Monte Carlo integration (see ~\ref{subsubsec:monte-carlo}).

\subsubsection{Spherical Integration}

To solve the integral numerically, we often parameterize the hemisphere $\Omega$
using spherical coordinates $(\theta, \phi)$.
The differential solid angle $d\omega_i$ changes variables as follows:
\begin{equation}
    d\omega = \sin\theta d\theta d\phi\label{eq:omega-cov}
\end{equation}

Substituting this into the rendering equation gives the explicit double integral form:

\begin{align}
    \int_{\Omega} (\dots) d\omega_i &= \int_{0}^{2\pi} \int_{0}^{\frac{\pi}{2}} L_i(x, \theta, \phi) \cdot \notag \\
    & f_r(x, \theta, \phi, \omega_o) \cos\theta \sin\theta d\theta d\phi
\end{align}

Here, $\theta$ is the polar angle (from the normal $n$, range $[0, \pi/2]$)
and $\phi$ is the azimuth angle (range $[0, 2\pi]$).
The term $\cos\theta$ comes from the geometry term $(\omega_i \cdot n)$,
and $\sin\theta$ comes from the differential solid angle measure.

\subsubsection{Color Spectrum}

Radiance $L$ varies with wavelength ($L(\lambda)$).
In computer graphics, strictly accurate rendering would require integrating over the entire visible spectrum.
However, for efficiency, we typically sample this spectrum at three discrete wavelengths corresponding to Red, Green,
and Blue (RGB).
This mimics the tristimulus response of the human eye.
While sufficient for most purposes, this approximation can fail for phenomena like
dispersion, fluorescence or metamerism result in inaccurate results.

\subsubsection{Energy Conservation (White Furnace Test)}

A physically-based material must conserve energy: it cannot reflect more light than it receives.
The ``White Furnace Test'' is a thought experiment (or unit-test) used to verify this.
If a non-absorbing object is placed in an environment with uniform radiance $L_i = 1$ everywhere,
the outgoing radiance $L_o$ must also be exactly $1$ (as long as no other form of light is emitted).
For a perfectly diffuse (Lambertian) surface with albedo $\rho$, the \gls{brdf} is constant $f_r = c$.
Integrating the cosine term over the hemisphere yields $\pi$:
\[
    \int_{\Omega} \cos\theta d\omega = \pi
\]
Therefore, to ensure $L_o = \rho L_i$, we must normalize the \gls{brdf}:
\[
f_{r, \text{diffuse}} = \frac{\rho}{\pi}
\]
The factor of $1/\pi$ is essential for energy conservation.