%%%%%%%%%%%%%%%%%%%%% chapter.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample chapter
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer Nature %%%%%%%%%%%%%%%%%%%%%%%%%%
\motto{The energy or active exercise of the mind constitutes life. ---Aristotle}
\chapter{Energy}
\label{energy}

\abstract*{
  Energy is all that matters in the universe. Every single planetary object, 
  every life form and every bit of dust is essentially a bunch of energy.
  Which kind of energy will be discussed in this chapter.
  For the purpose of this book, the energy momentum tensor will be introduced 
  and be the main focus of this chapter, while there will be some other 
  tensors involved as well.
}

\abstract{
  Energy is all that matters in the universe. Every single planetary object, 
  every life form and every bit of dust is essentially a bunch of energy.
  Which kind of energy will be discussed in this chapter.
  For the purpose of this book, the energy momentum tensor will be introduced 
  and be the main focus of this chapter, while there will be some other 
  tensors involved as well.
}

\section{Forms of Energy}
\label{sec:energy-forms}

\section{Conservation of Energy}
\label{sec:conservation-energy}

\section{Energy Momentum Tensor}
\label{sec:energy-momentum-tensor}

\subsection{Perfect Fluids}
\label{subsec:t-fluids}

\section{Einstein's conjecture}
\label{sec:eep}

\subsection{Weak Equivalence Principle}
\label{subsec:wep}

\subsection{Einstein Equivalence Principle}
\label{subsec:eep}

\subsection{Strong Equivalence Principle}
\label{subsec:sep}

\section{Torsion Tensor}
\label{sec:torsion-tensor}

\section{Weyl Tensor}
\label{sec:weyl-tensor}

\section{Cosmological Constant}
\label{sec:cosmos-constant}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% References
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{spmpsci}
\bibliography{ref-energy}
