%%%%%%%%%%%%%%%%%%%%% chapter.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample chapter
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer Nature %%%%%%%%%%%%%%%%%%%%%%%%%%
\motto{Imagination is more important than knowledge. ---Albert Einstein}
\chapter{Physics Introduction}
\label{physics}

\abstract*{
  Physics is mostly turned into cosmology, when diving deep into subjects
  covered by this book. But that does not exclude purely physics related 
  topics from appearing in this book. For this purpose this chapter includes
  sections about the basics of thermodynamics, the general appearance of 
  symmetries in the universe, angular momentum, the Doppler effect, Lagrangian
  mechanics and, of course, gravitation.
  Gravitation in itself is a complex concept but in this chapter only the 
  essentials in classical mechanics will be covered. 
}

\abstract{
  Physics is mostly turned into cosmology, when diving deep into subjects
  covered by this book. But that does not exclude purely physics related 
  topics from appearing in this book. For this purpose this chapter includes
  sections about the basics of thermodynamics, the general appearance of 
  symmetries in the universe, angular momentum, the Doppler effect, Lagrangian
  mechanics and, of course, gravitation.
  Gravitation in itself is a complex concept but in this chapter only the 
  essentials in classical mechanics will be covered. 
}

\section{Thermodynamics}
\label{sec:thermodynamics}

\subsection{The laws of thermodynamics}
\label{subsec:laws}

\subsection{The notion of temperature}
\label{subsec:temperature}

\subsection{Heat Transfer}
\label{subsec:heat-transfer}

\section{Symmetries in the universe}
\label{sec:symmetries}

\subsection{Noether Theorem}
\label{subsec:noether-theorem}

\subsection{Conservation Laws}
\label{subsec:conservation}

\section{Angular Momentum}
\label{sec:angular-momentum}

\subsection{General formulation}
\label{subsec:formulation-angular-momentum}

\subsection{Specific Angular Momentum}
\label{subsec:specific-angular-momentum}

\section{Lagrangian Mechanics}
\label{sec:lagrangian-mechanics}

\section{The Doppler Effect}
\label{sec:doppler}

\subsection{Sound Waves}
\label{subsec:sound-waves}

\subsection{Light Waves}
\label{subsec:light-waves}

\section{Gravitation}
\label{sec:gravitation}

\subsection{Newton's formulation}
\label{subsec:gravity-newton}

\subsection{Poisson's equation}
\label{subsec:gravity-poisson}

\subsection{Gravitational Potential}
\label{subsec:gravitational-potential}

\subsection{Field Theory}
\label{subsec:field-theory}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% References
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{spmpsci}
\bibliography{}
